All just Greedy for Money? - A prediction and Analysis of the interplay between Graduate Statistics and Salary Prospects in the STEM Field

### 2024-01-29 doc/DataLiteracyStudentProject/paper.tex
latex formatting

\begin{figure*}
    \includegraphics{fig/prediction_5.pdf}
    \caption{VAR model Prediction with optimal lag value of 5 on the test data set.}
\end{figure*}

@@

\subsection{Choosing lag}\label{sec:methods:lag}
To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria. Particularly the AIC and 
BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
We choose $\alpha = 5\%$ and get get a 95\% Confidence Interval 
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates \\
        \hline
        Prediciton 1 & 1.257\% & 1.271\% & 3.944\% \\
        \hline
        Prediciton 2 & 4.843\% & 0.119\% & 1.722\%
    \end{tabular}
    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to 1.}
\end{table}
The percentage differences between the true values especially for the students are quite high. To get better results
the AIC and BIC are used to find a better lag for the hyperparameter $p$ \textcolor{red}{is $p$ = lag?}. The abbreviation AIC
stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data.
This should prevent overfitting. 
BIC is short for Bayesian Information Criterion and penalizes more complex models.
Therefore BIC leads to less complicated models than the ones from AIC, which sets the two apart.
The optimal lag based on AIC is 6, while for BIC it is 1, which is consistent with the PACF.
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates \\
        \hline
        Prediciton 1 & 0.484\% & 0.419\% & 0.111\% \\
        \hline
        Prediciton 2 & 1.346\% & 0.306\% & 0.939\%
    \end{tabular}
    \caption{Percentage difference between the true value and predicted value for students, salary, and graduates. Lag set to 5.}
\end{table}

The percentage errors show that there is a significant improvement in the prediction of student numbers compared to the lag order of 1.
Setting lag to 5 results in even better errors for the domains of our interest, predicting students.

@@

% I am currently here
% Here the explanation for the correlation matrix
% Here teh the explanation for the visualisation of coefficients

With the correct lag chosen, the model fits the data by minimizing the mean square error.
The returned coefficients describe the impact of a specific lagged value from one time series
on predicting the next value. \ref{fig:student coefficients} visualizes all coefficients
relevant for predicting the number of students in the next semester. 
Notably, salary is the most relevant information for the prediction. All 5
coefficient values for each time step are non-zero, indicating a strong correlation.
Graduates, on the other hand, are not as relevant, recognizable by values close to zero.

@@

\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates\\
        \hline
        students & 1.000 & -0.115 &  0.111\\
        \hline
        salary & -0.115 & 1.000000 & -0.902\\
        \hline
        graduates & 0.111 & -0.902 &  1.000\\
    \end{tabular}
    \caption{Correlation matrix of residuals using 5 lags.}
    \label{table:correlation matrix}
\end{table}

>>>>

\begin{figure*}
    \includegraphics{fig/prediction_5.pdf}
    \caption{VAR model Prediction with optimal lag value of $5$ on the test data set.}
\end{figure*}

@@

\subsection{Choosing lag}\label{sec:methods:lag}
To find the optimal hyperparameter we used two well-established methods, the PACF and information-based criteria. Particularly the AIC and 
BIC. The PACF only considers the impact of the previous data point $S_{n-1}$ on $S_{n}$ to build a $(1-\alpha)$ confidence interval.
We choose $\alpha = 5\%$ and get get a 95\% Confidence Interval 
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates \\
        \hline
        Prediciton 1 & 1.257\% & 1.271\% & 3.944\% \\
        \hline
        Prediciton 2 & 4.843\% & 0.119\% & 1.722\%
    \end{tabular}
    \caption{Difference in percent between the true value and predicted value for students, salary, and graduates. Lag set to 1.}
\end{table}
The percentage differences between the true values especially for the students are quite high. To get better results
the AIC and BIC are used to find a better lag for the hyperparameter $p$ \textcolor{red}{is $p$ = lag?}. The abbreviation AIC
stands for Akaike Information Criterion and tries to find an optimum between the complexity of the model and fitting the data.
This should prevent overfitting. 
BIC is short for Bayesian Information Criterion and penalizes more complex models.
Therefore BIC leads to less complicated models than the ones from AIC, which sets the two apart.
The optimal lag based on AIC is 6, while for BIC it is 1, which is consistent with the PACF.
\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates \\
        \hline
        Prediciton 1 & 0.484\% & 0.419\% & 0.111\% \\
        \hline
        Prediciton 2 & 1.346\% & 0.306\% & 0.939\%
    \end{tabular}
    \caption{Percentage difference between the true value and predicted value for students, salary, and graduates. Lag set to $5$.}
\end{table}

The percentage errors show that there is a significant improvement in the prediction of student numbers compared to the lag order of 1.
Setting lag to $5$ results in even better errors for the domains of our interest, predicting students.

@@

With the correct lag chosen, the model fits the data by minimizing the mean square error.
The returned coefficients describe the impact of a specific lagged value from one time series
on predicting the next value. \ref{fig:student coefficients} visualizes all coefficients
relevant for predicting the number of students in the next semester. 
Notably, salary is the most relevant information for the prediction. All five
coefficient values for each time step are non-zero, indicating a strong correlation.
Graduates, on the other hand, are not as relevant, recognizable by values close to zero.

@@

\begin{table}[H]
    \centering
    \begin{tabular}{c|c|c|c}
        & students & salary & graduates\\
        \hline
        students & 1.000 & -0.115 &  0.111\\
        \hline
        salary & -0.115 & 1.000000 & -0.902\\
        \hline
        graduates & 0.111 & -0.902 &  1.000\\
    \end{tabular}
    \caption{Correlation matrix of residuals using $5$ lags.}
    \label{table:correlation matrix}
\end{table}
